---
title: "Exploratory Data Analysis"
output: github_document
---

```{r setup}
library(tidyverse)
library(patchwork)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 0.6,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom")) 

options(
  ggplot2.continuous.color = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d()
scale_fill_discrete = scale_fill_viridis_d()
```

Pull RNOAA data.

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, everything())

weather_df
```

## `group_by`

Could group_by within a location value and set of month values 

can also `ungroup()`

```{r}
weather_df %>% 
  group_by(name, month)
```

## Counting things

Count month / name observations

```{r}
weather_df %>% 
  group_by(name) %>% 
  summarize(n_obs = n())
```

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(n_obs = n())
```

We can use `count()` as well

```{r}
weather_df %>% 
  count(name, month, name = "n_obs")
```

**NEVER** use base R's `table`. Result is not a dataframe. Never pull variables out of dataframe.

```{r eval = FALSE}
weather_df %>% 
  pull(month) %>% 
  table()
```

Other helpful counters

```{r}
weather_df %>% 
  group_by(month) %>% 
  summarize(
    n_obs = n(),
    n_days = n_distinct(date))
```

In summarize, created 2 variables, number observations and number of distinct dates in that month.

## A digression on 2x2 tables

Define: name(Central Park and Waterhold) as exposure, and variable on whether that date is cold or not cold

```{r}
weather_df %>% 
  filter(name != "Waikiki_HA") %>% 
  mutate(
    cold = case_when(
      tmax < 5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE      ~ ""
    )
  ) %>% 
  group_by(name, cold) %>% 
  summarize(count = n())
```

Quicker way to do this.

```{r}
weather_df %>% 
  filter(name != "Waikiki_HA") %>% 
  mutate(
    cold = case_when(
      tmax < 5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE      ~ ""
    )
  ) %>% 
  janitor::tabyl(name,cold)
```

## General summaries

You can do lots of summaries.

```{r}
weather_df %>% 
  group_by(month) %>% 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE),
    mean_prcp = mean(prcp, na.rm = TRUE),
    median_tmin = median(tmin, na.rm = TRUE)
  )
```

This is a dataframe!

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE),
    mean_prcp = mean(prcp, na.rm = TRUE),
    median_tmin = median(tmin, na.rm = TRUE)
  ) %>% 
  ggplot(aes(x = month, y = mean_tmax, color = name)) +
  geom_point() +
  geom_line()
```

This is similar to

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_line()
```

Suppose you want to summarize many columns. If apply same summarize function to multiple variables using `across` within `summarize()`

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(across(prcp:tmin, mean))
```

Reminder: sometimes your results are easier to read in another format.

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(mean_tmax = mean(tmax)) %>% 
  pivot_wider(
    names_from = name,
    values_from = mean_tmax
  ) %>% 
  knitr::kable(digits = 1)
```


