---
title: "Visualization 1"
output: github_document
---

```{r setup}
library(tidyverse)
library(ggridges)
```

Pull RNOAA data.

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df
```

## Create scatterplots

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```

New approach, same plot. Could use this %>% argument scheme to add in other functions like mutate. 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

Save and edit a plot object. Can then later transform and print in other ways

```{r}
weather_plot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax))

weather_plot + geom_point()
```

## Advanced scatterplot

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE)
```

What about the `aes` placement?

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name)) +
  geom_smooth()
```

This color aesthetic only applies to scatterplot, so the geom_smooth doesn't differentiate by color. 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

facet_grid: "." nothing defines rows "~ names" means names defines columns. Here's the other way around.

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  facet_grid(name ~ .)
```

alpha in geom_point as transparency. 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.2, size = 1) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

Size of point and transparency could be defined based on a variable.

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, alpha = tmin, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

Combine some elements and try a new plot. 

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = 0.5) + 
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

## Some small notes

How many geoms have to exist? Can use whatever geoms I want. Can go straight to geom_smooth if want. 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_smooth(se = FALSE)
```

You can use a neat geom!

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_density2d() + 
  geom_point(alpha = 0.3)
```

## Univariate plot

Histogram - helpful to understand what's going on in dataset

```{r}
weather_df %>% 
  ggplot(aes(x = tmin)) + 
  geom_histogram()
```

Can we add color. Overlayed distributions

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) + 
  geom_histogram(position = "dodge")
```

Separate distribution

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) + 
  geom_histogram() + 
  facet_grid(. ~ name)
```

If want to compare distribution across levels (or location, categorical var), can do density plot, which is like histogram with smoothed edges.

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = 0.4)
```

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = 0.4, adjust = 0.5)
```

Boxplots! Can use categorical var on x axis, and then y var is the continuos var.

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmin)) + 
  geom_boxplot()
```

Trendy plots

violin plots - cross between density plot and boxplot

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_violin(alpha = 0.5) + 
  stat_summary(fun = "median")
```

Ridge plots - density curve. Particularly good for categorical independent predictor variable with large number of values (like the 50 states) and visualizing categorical var on that.

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = name)) +
  geom_density_ridges()
```

## Save and embed

```{r}
weather_plot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.5)

ggsave("./results/weather_plot.pdf", weather_plot, width = 8, height = 5)
```

What about embedding. including specifying different size

```{r, fig.width=12, fig.height=6}
weather_plot
```

```{r, fig.width=6, fig.height=3}
weather_plot
```

