---
title: "Visualization 2"
output: github_document
---

```{r setup}
library(tidyverse)
library(patchwork)
```

Pull RNOAA data.

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df
```

## Starting plot

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5)
```

## Labels

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017"
  )
```

## Scales

Scale_x_continuous and scale_x_discrete
Scale_y_continuous as "sqrt" or "log"

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017"
  ) + 
  scale_x_continuous(
    breaks = c(-15, 0, 15),
    labels = c("-15 C", "0 C", "15 C")
  ) + 
  scale_y_continuous(
    trans = "log"
  )
```

Look at color scales.

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017"
  ) +
  scale_color_hue(
    name = "location",
    h = c(100, 300))
```

Check out viridis color palatte. Can use magma, plasma, inferno. 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017"
  ) +
  viridis::scale_color_viridis(
    name = "location",
    discrete = TRUE
  )
```

## Themes

Specifications that aren't data dependent. 
Shift legend. 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017"
  ) +
  viridis::scale_color_viridis(
    name = "location",
    discrete = TRUE
  ) + 
  theme(legend.position = "bottom")
```

Change the overall theme. Search ggthemes

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017"
  ) +
  viridis::scale_color_viridis(
    name = "location",
    discrete = TRUE
  ) + 
  theme_bw()
```

Need general overall theme argument before specific argument like legend location 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017"
  ) +
  viridis::scale_color_viridis(
    name = "location",
    discrete = TRUE
  ) + 
  theme_minimal() + 
  theme(legend.position = "bottom")
```

## Setting options

```{r, eval = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 0.6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom")) 

options(
  ggplot2.continuous.color = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_color_discrete = scale_color_viridis_d()
scale_fill_discrete = scale_fill_viridis_d()
```

## Data args i 'geom'

Can have one geometry with one dataset and another geom for another dataset. Create two small datasets and use data argument to determine which dataset is relevant to which geometry.

```{r}
central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) +
  geom_point() + 
  geom_line(data = central_park)
```

## Patchwork

Faceting creates same geometry of plot separated by variable name

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = 0.5) +
  facet_grid(. ~ name)
```

If want to put multi-panel plots of different geometry (can't facet). Make individual plots and patch together. 

```{r}
tmax_tmin_p = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.5) +
  theme(legend.position = "none")

prcp_dens_p = 
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = 0.5)

tmax_date_p = 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() + 
  geom_smooth(se = FALSE) + 
  theme(legend.position = "none")

tmax_tmin_p + prcp_dens_p + tmax_date_p
(tmax_tmin_p + prcp_dens_p) / tmax_date_p
```

